---
title: "Course project 1"
#Weiye Deng
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. reading in the dataset and/or processing the data
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "repdata_data_activity.zip",mode="wb")
unzip("repdata_data_activity.zip")
activity <- read.csv("activity.csv",header=TRUE)
head(activity,30)
tail(activity,30)
```

## 2. Histogram of the total number of steps taken each day
```{r}
## total number of steps taken each day
total_num <- with(activity, tapply(steps, as.factor(activity$date), sum, na.rm = T))
hist(total_num, main="Histogram of the total number of steps taken each day", xlab="Total number of steps", ylab="Frequemcy (%)")
```


## 3. Mean and median number of steps taken each day
```{r}
summary(total_num)
```
#mean and median values are 10766.

## 4. Time series plot of the average number of steps taken
```{r}
mean_steps <- aggregate(steps ~ interval, activity, mean)

plot(mean_steps$interval, mean_steps$steps, type='l', col=1, 
     main="Time series plot of the average number of steps taken", xlab="Interval", 
     ylab="Average number of steps")

```



## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max_num <- which.max(mean_steps$steps)
mean_steps [max_num, ]

```
#The interval is 835.

## 6.Code to describe and show a strategy for imputing missing data
```{r}
#missing data
sum(is.na(activity))
missing_row <- activity[is.na(activity$steps),]

#replace NA with neab values of 5-minute interval
for (i in 1:nrow(activity)) {
    if(is.na(activity$steps[i])) {
        val <- mean_steps$steps[which(mean_steps$interval == activity$interval[i])]
        activity$steps[i] <- val 
    }
}

impute_dt <- aggregate(steps ~ date, activity, sum)


```
#2304 row missing data.



## 7.Histogram of the total number of steps taken each day after missing values are imputed
```{r}

hist(impute_dt$steps, main = "Histogram of total number of steps per day (Imputation)", xlab = "Steps per day")

```




## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
#define weekend and weekday variables
day <- weekdays(as.Date(activity$date))

two_grps <- vector()
for (i in 1:nrow(activity)) {
    if (day[i] == "Saturday" || day[i] == "Sunday"){
        two_grps[i] <- "Weekend" }
     else {
        two_grps[i] <- "Weekday"
    }
}


activity$two_grps <- two_grps
activity$two_grps <- factor(activity$two_grps)

stepsByDay <- aggregate(steps ~ interval + two_grps, data = activity, mean)
names(stepsByDay) <- c("interval", "two_grps", "steps")



library(lattice)
xyplot(steps ~ interval | two_grps, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")


```
#Difference was observed in the figures between weekend and weekday. Users at weekdays had earlier interval and more step numbers compared with users at weekend.




















